Refinements for Open Automata

### 2023-06-23 2021-Refinement/paper/refinementarticleV2.tex
correction

\RAB{Skeleton-based systems, such as skeleton-based parallel programs, consist of  a collection of high-level patterns with parallel semantics,  but also provide a systematic way of developing practical programs. Skeleton specification is refined by applying a sequence of transformation rules until implementation.  In this high-level parallel design  approach, each skeleton is a parametric process, it accepts as a parameter the kind of computation to be performed according to the behaviour it models.
}

@@

There are not a lot of works that study refinement for open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
\cite{GayH:2005,BravettiZ:2021}. This kind of refinement is  based  on an interface-oriented approach, it allows the expression of more\LUDO{more and lesss than what???} internal choices and less external choices. The refinement of open systems is also defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}.
Alternating simulation  is originating from the game theory \cite{deAlfaro:2003}, it allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement of game-based automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions). However, the composition of such automata may
lead to illegal states, where one automaton issues an output that is not acceptable as input in the other one. The theory of alternating simulation provides an optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, two components can be composed if there is an environment where they can work together. Our approach  has some commonalities with  the above mentioned simulation \cite{deAlfaro:2021}: both are process-oriented approaches even if they are not based on the same notion of simulation and both are based on optimistic approach to composition. Rather than being prepared to cope with all environments, components are described together with the environment with which they will interact. \LUDO{\st{we should explain that we are optimistic, then}} \LUDO{Je comprends pas l argument: ca a l air de dire qu on fait pas des systemes ouverts. en effet dans ce papier on decrit pas l environnement dans ce travail, a ce que je comprends? A REVOIR}. They 
For the composability, we use the notion of comparability of holes \LUDO{do we? where?} (similar to the notion of compatibility), which is explicitly encompassed in the definition of composition.

@@

The simulation  ensures the preservation of   
safety properties as deadlock-freeness  and, more generally, 
all linear temporal logic properties \cite{Manna:1992} \LUDO{\st{add ref!}} \LUDO{add ref sur ``simulation  ensures the preservation of  safety properties'' pas sur temporal logic!}.  Our focus is about the properties of simulation, in particular closure properties such as preservation under composition. Some works   have addressed the problem of non-introduction of blocking by composition. For example,  \cite{DIHEGO2020110598}  presented a refinement relation on process algebra based on failure semantics that is basically different from the (bi)simulation \cite{10.5555/640428.640430}.   \LUDO{the last sentence is not really informative} \LUDO{how do we compare our work?}

>>>>

\RAB{Many parallel programs  such as skeleton-based parallel programs,  can be characterized  by a collection of high-level patterns with parallel semantics,  but also provide a systematic way of developing practical programs. Skeleton specification is refined by applying a sequence of transformation rules until implementation.  In this high-level parallel design  approach, each skeleton is a parametric process, it accepts as a parameter the kind of computation to be performed according to the behaviour it models.
}

@@

There are not a lot of works that study refinement for open systems. Defining refinement of open systems as trace inclusion  is  addressed  as a notion of subtyping in type theory 
(e.g. \cite{GayH:2005,BravettiZ:2021}). The definition of refinement is based on a connection  between session types and communicating automata theories -- a notion of session automata
based on Communicating Finite-State Machines, that are used for  modelling processes communicating through FIFO channels.   \LUDO{\st{more and lesss than what???}}. The refinement of open systems is also defined in terms of  alternating simulation \cite{Alur:1998,deAlfaro:2021}.
Alternating simulation  is originating from the game theory \cite{deAlfaro:2003}, it allows  the study of relation between individual components by viewing them as alternating transition systems. In particular,  a refinement of game-based automata expresses that the refined component can offer more services (input actions) and fewer service demands (output actions). However, the composition of such automata may
lead to illegal states, where one automaton issues an output that is not acceptable as input in the other one. The theory of alternating simulation provides an optimistic approach to compute compatibility between automata based on the fact that each automaton expects the other to provide  legal inputs, i.e, two components can be composed if there is an environment where they can work together. Our approach  has some commonalities with  the above mentioned simulation \cite{deAlfaro:2021}: both are process-oriented approaches even if they are not based on the same notion of simulation and both are based on optimistic approach to composition. Rather than being prepared to cope with all environments, components are described together with the environment with which they will interact. \LUDO{\st{we should explain that we are optimistic, then}} \LUDO{Je comprends pas l argument: ca a l air de dire qu on fait pas des systemes ouverts. en effet dans ce papier on decrit pas l environnement dans ce travail, a ce que je comprends? A REVOIR}. They 
For the composability, we use the notion of comparability of holes \LUDO{do we? where?} (similar to the notion of compatibility), which is explicitly encompassed in the definition of composition.

@@

The simulation  ensures the preservation of   
safety properties as deadlock-freeness  and, more generally, 
all linear temporal logic properties in the sense of \cite{Abrialbooks96,Kouchnarenko:2007} \LUDO{\st{add ref sur ``simulation  ensures the preservation of  safety properties'' pas sur temporal logic!}}.  Our focus is about the properties of simulation, in particular closure properties such as preservation under composition. Some works   have addressed the problem of non-introduction of blocking by composition. For example,  \cite{DIHEGO2020110598}  presented a refinement relation on process algebra based on failure semantics that is basically different from the (bi)simulation \cite{10.5555/640428.640430}.   \LUDO{the last sentence is not really informative} \LUDO{how do we compare our work?}
